### 2010-05-16 Z0/tex/quellen.tex

\renewcommand{\refname}{Literatur und Programme}
\begin{thebibliography}{xxxxxx}
\bibitem[Skript]{script}
M. zur Nedden, Anleitung zum Versuch $Z^0$–Resonanz, Humboldt–Universität, 2008
\bibitem[PDB]{pdb}
Particle Data Group, Particle Physics Booklet, Physics Letters, 2008
\bibitem[LEP91]{lep91}
H. Schopper, Die Ernte nach einem Jahr LEP–Betrieb, Physikalische Blätter, 1991
\bibitem[ADR]{kap5}
O. Adrinai et al., Results from the L3 Experiment at LEP,  Physics Reports, 1993
\bibitem[ROOT]{root}
ROOT Version 5.26/00, 2009, \href{http://root.cern.ch}{http://root.cern.ch}
\end{thebibliography}