### 2010-08-01 Z0-Publikation/tex/paper-template.tex
weiter

%%%%%%%%%%%%%%%%%%%%%%%%% file draft4_2.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	                                                                   %
% Based on the  template file for The European Physical Journal C          %
%                                                                          %
%                                                                          %
%                                                                          %
%                                                                          % 
%%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[epj,nopacs]{svjour}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{../../assets/uniinput}

\usepackage{siunitx}
\usepackage[justification=raggedright,singlelinecheck=false]{caption}
\sisetup{
	seperr		=	true,
	trapambigerr	=	true,
	openerr		=	(,
	closeerr	=	),
	expproduct	=	cdot,
	padnumber	=	both,
	stickyper	=	true,
	per		=	slash,
	trapambigfrac	=	true,
	repeatunits	=	false,
	openfrac	=	(,
	closefrac	=	),
	prefixsymbolic 	=	true,
	prefixproduct	=	cdot,
	decimalsymbol	=	.,
        tabnumalign     =       left,
        tabtextalign    =       left
}

\newcommand{\spar}{{\stackrel{\rightarrow}{\Rightarrow}}}
\newcommand{\sant}{{\stackrel{\rightarrow}{\Leftarrow}}}
\newcommand{\sperpa}{{\rightarrow\Uparrow}}
\newcommand{\sperpb}{{\rightarrow\Downarrow}}
\begin{document}
\hugehead

\newcommand{\dd}[1]{\mathrm{d}#1\,} % declare dx operator, usage: \dd{x} for dx
\newcommand{\lref}[1]{listing (\ref{lst:#1})} % refer to a listing, usage:\lref{label} for Listing (...)
\newcommand{\fref}[1]{fig. (\ref{fig:#1})} % refer to a figure, usage:\lref{label} for Abb. (...)
\newcommand{\tref}[1]{tab. (\ref{tab:#1})} % refer to a table, usage:tlref{label} for Tab. (...)
\newcommand{\eref}[1]{eq. (\ref{eqn:#1})} % refer to a equation, usage:\eref{label} for Gl. (...)

\title{Z-Resonance at LEP}
\titlerunning{Z-Resonance at LEP}
\authorrunning{Thomas Murach, Robert Riemann}
\author{{\bf Draft Version 1.1}\\
\medskip \\
Thomas Murach,$^{1}$
Robert Riemann$^{1}$
} 
\institute{$^1$Department of Physics, Humboldt University of Berlin, Germany\\}  
\date{Received: \today / Revised version:}
\abstract{
We had a lot of fun measuring the lifetime and the mass of the Z boson! These quantities 
are important to answer the question of life, universe and everything.
Another nice thing we could access was the number of the lepton families, which we 
were especially happy about...  
}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-1.5cm}
\section{ Introduction}
\baselineskip=0.38cm
\vspace*{1.cm}

The standard model of elementary particles describes interactions of particles
by four fundamental forces. One of them is the electroweak force. Its gauge
bosons, which mediate the force between the two interacting particles, are the
photon as well as the W$^{\pm}$- and the Z$^0$-boson. At the
electron-positron-collider LEP at CERN the Z$^0$ has been seen for the first
time, which was possible because of the sufficient center of mass energy of this
collider experiment.

We were able to evaluate its mass $M_Z$ and its lifetime τ. The leptonic and
the hadronic decay widths $Γ_l$ and $Γ_h$ could be investigated as well. Finally
the number of quark and lepton families $N_C$ and the weak mixing angle, also
called Weinstein angle, $\Theta_W$ can be measured. This angle describes the
ratio of the electromagnetic and the weak coupling constants α and α$_w$.

\section{ Experiment}

LEP was designed to be able to deliver center of mass energies $\sqrt{s}$ from
some 89 up to \SI{209}{\giga\electronvolt}. This is enough to produce Z-bosons,
which should have a mass of about \SI{91}{\giga\electronvolt}, as it was
predicted by the standard model. Therefore the Z$^0$-particle was supposed to be
produced at LEP.

It can decay in various ways, but always into pairs of one
fermion and one antifermion. One of these possibilities is the decay via two
different quarks (Z$^0→q\overline{q'}$), another one could be the decay via two
charged leptons like Z$^0→ μ^+μ^-$. The first one of the mentioned decays can be
seen as two jets which can be measured in hadronic calorimeters. The leptonic
decay is measured due to registration of leptons in electromagnetic
calorimeters. All of these layers have an almost cylindrical shape in order to
cover as much of the solid angle as possible. This is also the reason for the
installation of the end caps.

The detector for all of these particles is LEP. Its structure can be seen
in \fref{aufbau}.
\begin{figure}[htb]
 \centering
 \includegraphics[page=5,viewport=286 620 508 765,clip,%
  width=\columnwidth,keepaspectratio]{../../Z0/docs/Z0ResFprakt}
 \caption{Design of the L3 detector \cite{script}}
 \label{fig:aufbau}
\end{figure}
There are layers of components with different purposes, starting from the
interaction point, these are the tracking chamber, the electromagnetic calorimeter,
the hadronic calorimeter and finally the muon chamber.

In this case the hadronic calorimeter and the muon chamber are the most
important parts. The hadronic calorimeter consists of alternating layers of
uranium plates and wire chambers. The muon chamber is built up of drift
chambers. It detects the high energy muons, which have passed the detector
without a significant loss of energy, because these are the only particles that
are able to reach there. Further information about the detector design can
be found in \cite{script}.

\section{ Data analysis}

For a detailed analysis of the mentioned processes it is important to be able
to distinguish between muonic and hadronic decays. By Monte Carlo studies the
difference between both processes could be figure out. This unlocks the
possibility to apply cuts on the measured distributions. in this paper first
results of measurements at center of mass energies of 89.48, 91.33 and
\SI{93.02}{\giga\electronvolt} can already be presented.

\subsection{ Event selection}
Muons barely interact with any matter, therefore there are only very few
entries in the calorimeter along the track of the muon. This has an immense
discriminating power, because jets consist of tens of particles due to
hadronization at high energies and the decay of unstable particles inside the
jets.

Due to inefficiencies in energy measurement it is possible that not all the
energy provided by the apparatus can finally be seen. Therefore there are large
uncertainties in jet energy measurements, whereas these uncertainties are
smaller for muonic events. This fact can also be used to distinguish these
events from each other and from background as well. For example if the $Z^0$
decays to two very unstable tauons, these tauons decay very fast. Then two
neutrinos are created, but they cannot be measured, so that their contribution
to the visible energy is missing. The visible part of the tau decays resembles
hadronic events, so it is important to be able to separate this background from
the hadronic signal.

To be able to separate muonic events from background it is suitable to ask for
an angle of almost π, as the $Z^0$ boson should be created almost resting in
the laboratory frame, leading to muons flying back to back. Another selection
for muonic events is possible through the requirement of both muons carrying
about half the energy of the event each. This is possible because muon energies
can be measured very well. Finally it is possible to recognize muons through
their mass.

\subsection{Hadronic events}
As it is now known how to extract hadronic events out of the data, it is
possible to determine the hadronic cross section using the following formula:
\begin{equation}
σ_{\mathrm{had}} = \frac{N}{Lε},
\end{equation}
where ε is the efficiency of the cuts. This efficiency could be calculated
using the following equation:
\begin{equation}
ε = \frac{N’_{\mathrm{MC}}}{N_{\mathrm{MC}}},
\end{equation}
where $N’_{\mathrm{MC}}$ is the number of Monte Carlo events that passed the
hadron cuts and $N_{\mathrm{MC}}$ is the total number of events in the Monte
Carlo file.

The hadronic cross section could be calculated at three different center of
mass energies. The number of hadronic events that passed the hadron cuts can be
found in \tref{hadr_events}, as well as the hadronic cross sections themselves.

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|l|}
\hline
energy $\sqrt{s}$ [\si{\GeV}] & hadronic events & cross section $σ_0^μ$ [\si{\nano\barn}]\\
\hline
89.48 & 1848 ± 43 ± 52 & 10.55 ± 0.18 ± 0.16 \\
91.33 & 3980 ± 63 ± 79 & 29.98 ± 0.49 ± 0.43 \\
93.02 & 2149 ± 46 ± 51 & 14.56 ± 0.24 ± 0.24 \\
\hline
\end{tabular}
\vspace*{0.3cm}
\caption{\baselineskip=0.38cm analyzed hadronic events at different center of mass energies}
\label{tab:hadr_events}
\end{center}
\vspace*{-0.5cm}
\end{table}

These cross sections have then been used as data for a fit. By fitting the data
points the overall hadronic cross section $σ_0^{\mathrm{had}}$, the mass $m_Z$ and the decay width $Γ_Z$
of the $Z^0$-boson could be determined. The results can be found in
\tref{results}.

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|l|}
\hline
$m_Z$ [\si{\GeV}] & $σ_0^{\mathrm{had}}$  [\si{\nano\barn}] & $Γ_Z$ [\si{\GeV}]\\
\hline
$91.16 \pm 0.02 \pm 0.02$ & $40.33 \pm 0.65 \pm 0.27$ & $2.61 \pm 0.05 \pm 0.03$\\
\hline
\end{tabular}
\vspace*{0.3cm}
\caption{\baselineskip=0.38cm results of fit on data: hadronic cross section $σ_0^{\mathrm{had}}$, the mass $m_Z$ and the decay width $Γ_Z$ of the $Z^0$-boson}
\label{tab:results}
\end{center}
\vspace*{-0.5cm}
\end{table}
This is the first and already very precise measurement of the mass of the $Z^0$-boson.

The χ$^2$-value of this fit is almost zero, but as there were only three points to fit, whereas at the same time three parameters had to be evaluated, the number of degrees of freedom (ndf) is zero. Therefore the value of $χ^2/$ndf is undefined. During the next data taking periods more data at different center of mass energies will be acquired, so that this problem will be solved.

From the decay width of the $Z^0$, it is also possible to derive the lifetime of this particle. This is done using the following formula:
\begin{eqnarray}
τ_Z &=& \frac{1}{Γ_Z}\\
&=& (2.53 \pm 0.31 \pm 0.02) \cdot \SI{e-25}{\second}
\end{eqnarray}

\section{muonic events}

On the other hand, muons could be distinguished from hadronic events as well as from background. Therefore an analysis on muonic events is possible. This is described in this section.

Firstly, the cross sections at the different center of mass energies could be calculated in an analogue way to the hadronic events. The results of this can be found in \tref{muon_events}.

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|l|}
\hline
energy $\sqrt{s}$ [\si{\GeV}] & muon events & cross section $σ_0^μ$ [\si{\nano\barn}]\\
\hline
89.48 & \phantom{0}67 ± \phantom{0}8 ± \phantom{0}8  & 0.64 ± 0.03 ± 0.3 \\
91.33 & 122 ± 11 ± 10 & 1.53 ± 0.05 ± 0.4 \\
93.02 & \phantom{0}77 ± \phantom{0}9 ± \phantom{0}9 & 0.87 ± 0.04 ± 0.03 \\
\hline
\end{tabular}
\vspace*{0.3cm}
\caption{\baselineskip=0.38cm analyzed muonic events at different center of mass
energies}
\label{tab:muon_events}
\end{center}
\vspace*{-0.5cm}
\end{table}

Starting from these values and already knowing the decay width of the $Z^0$-boson as well as its mass, we can again do a fit on these three data points. This time the only free parameter is the muonic cross section $σ_0^μ$.

The result of this fit is shown in the following line:
\begin{equation}
σ_0^μ = (2.16 \pm 0.05 \pm 0.03) \si{\nano\barn}
\end{equation}
The $χ^2$-value of this fit is some 6.8, which gives a $χ^2/$ndf of about 3.4.
This is already in the right order of magnitude (close to 1).

\subsection{Background subtraction}

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\multicolumn{1}{|c|}{}& proton & deuteron\\
\hline
\multicolumn{3}{|c|}{Exclusive electroproduction}\\
\hline
$A^\rho_1$       & 0.23 $\pm$ 0.13$\pm$0.01 & -0.040 $\pm$ 0.076$\pm$0.003\\ 
$A^\phi_1$       & 0.20$\pm$0.45$\pm$0.01  & 0.17$\pm$0.27$\pm$0.01\\
$N^\rho$         &1774                   & 6505\\
$N^\phi$         & 219                   & 618 \\
\hline
\multicolumn{3}{|c|}{Electroproduction by quasi-real photons}\\
\hline
$A^\rho_1$  & 0.0057 $\pm$ 0.0093$\pm$0.0004  & -0.0039 $\pm$ 0.0029$\pm$0.0003\\ 
$A^\phi_1$  & 0.052 $\pm$ 0.084$\pm$0.003  & 0.018 $\pm$ 0.028$\pm$0.001\\
$N^\rho$    & $423\times10^{3}$ & $4013\times10^{3}$\\
$N^\phi$ & $7.6\times10^{3}$ &  $57\times10^{3}$ \\
\hline
\end{tabular}
\vspace*{0.3cm}
\caption{\baselineskip=0.38cm Example of a table}
\label{tab_asy}
\end{center}
\vspace*{-0.5cm}
\end{table}

\section{ Summary}

\begin{thebibliography}{}
\baselineskip=0.38cm
\bibitem{z1} T. H. Bauer et al., Rev. Mod. Phys. \textbf{50} No 2 (1978) 261  
\bibitem{z2}D. Schildknecht et al., Phys. Lett. \textbf{B 449} (1999) 328
\bibitem{z3}A. Donnachie, P. V. Landshoff, Phys. Lett. \textbf{B 478} (2000) 146

\end{thebibliography}

\begin{figure}[htb]
 \centering
 \includegraphics[width=1\columnwidth,keepaspectratio]{E_vis}
 \caption{Distribution of visible energie}
 \label{fig:e_vis}
\end{figure}

\begin{figure}[htb]
 \centering
 \includegraphics[width=1\columnwidth,keepaspectratio]{N_Cluster}
 \caption{Distribution of entries in the calorimetre}
 \label{fig:n_cluster}
\end{figure}

\begin{figure}[htb]
 \centering
 \includegraphics[width=1\columnwidth,keepaspectratio]{finalhad_fit}
 \caption{Breit-Wigner-Fit hadrons}
 \label{fig:e_vis}
\end{figure}

\begin{figure}[htb]
 \centering
 \includegraphics[width=1\columnwidth,keepaspectratio]{finalmu_fit}
 \caption{Breit-Wigner-Fit muons}
 \label{fig:n_cluster}
\end{figure}

\end{document}