### 2010-10-24 Quantenkryptographie/tex/anhang.tex
versuchsziel eigentl. fertig, anhang vorbereitet

\appendix
\section{Wahl der Polarisation}
\label{sec:zirkular}
Text.