### 2010-10-24 Quantenkryptographie/tex/anhang.tex
messwerte EOM drin, Anhang Bild

\appendix
\section{Abschwächung des Lasers}
\label{sec:laser}
Text?

\section{Wahl der Polarisation}
\label{sec:zirkular}
Text.

\onecolumn
\section{Messprotokoll}
\label{sec:protokoll}
\begin{figure}[!ht]
        \centering
        \includegraphics[page=1,width=.88\textwidth,keepaspectratio]{../data/messprotokoll}
%         \caption{Messprotokoll}
        \label{fig:protokoll}
\end{figure}